---
title: "In-Class Exercises #1"
subtitle: "SDS 270: Programming for Data Science in R, Spring 2026"
author: "YOUR NAME HERE"
editor: source
date: last-modified
---

## Introduction

These exercises are two-fold: to get you familiar with the layout of these assignments and to review functionality of (atomic) vectors in R.

There are two types of vectors in R: lists and atomic vectors. For now, we shift our focus to atomic vectors, which are defined by having the same type for all elements within them.

This means that if we attempt to create a simple vector of characters and doubles, such as:

```{r}
test <- c("a", 1, "c", 2)
```

R will automatically coerce the least-flexible types in the vector to the most flexible, meaning that our `test` vector will consist of solely character types. The exercises below will help you develop a fundamental breakdown of atomic vectors and their usage. To do so, we will analyze a simple dataframe.

## Diamonds Data Frame

We will look at the `starwars` dataframe from the `dplyr` package, which contains characters from the first two trilogies of Star Wars movies. Use the code below to load the `starwars` dataframe in:

```{r}
library(dplyr)
data(starwars)
```

## Questions

::: {.callout-important}

## Question 1

Provide at least two reasons why we would want to have a data structure that strictly contains elements of the same type. 

:::

::: {.callout-important}

## Question 2

Extract the `height` column from the `starwars` dataframe with `starwars$height` and assign it to `height`. Determine its type and whether it is a valid type given what the column stands for.

:::

::: {.callout-important}

## Question 3

Perform three operations using the `height` and `mass` columns from the `starwars` dataframe.

:::

::: {.callout-important}

## Question 4

For each R type (i.e. logical, integer, character, double), provide examples of real-world data that would be best suited to that R type and justify your choice.

:::

::: {.callout-important}

## Question 5

Create an R object of an atomic vector of at least 5 elements based on what was discussed in Q4. Then, add an attribute of your choice to the vector.

:::

When finished, commit and push this file to your SDS 270 repository.

